Short and Mechanized Logical Relation for Dependent Type Theory

### 2024-02-22 paper/paper.tex
Minor wording

For the other direction, given a function $[[F]]$ satisfying the
property $[[forall a, (# a in S implies Interp I i B { a } F a #) ]]$, it
is possible to go in the other direction and define a relation $[[R]]=[[
{ ( a , F a )  | a in S }]]$. It is easy to verify that $[[R]]$ satisfy
the the constraint required in the precondition of \rref{I-Pi}.

@@

Unfortunately, \rref{I-PiAlt} not only violates the syntactic strict
positivity constraint required in proof assistants, but
is genuinely non-monotone when written as an endofunction over the
domain of relations.
Intuitively, the failure of monotonicity stems from the fact
that the witness picked in the precondition is not necessarily the
same witness being referred to in the post condition. While it might
be possible to restrict the domain with additional constraints such as
functionality and inversion properties, we opt for our current
formulation of \rref{I-Pi} so we immediately obtain a
well-defined inductive relation and a usable induction principle. The
slight disadvantage of \rref{I-Pi} is that we need to construct the
function $[[R]]$ each time we apply it, though this is mitigated by
the immediate admissibility of \rref{I-PiAlt}.

>>>>

For the other direction, given a function $[[F]]$ satisfying the
property $[[forall a, (# a in S implies Interp I i B { a } F a #) ]]$, it
is possible to  define a relation $[[R]]=[[
{ ( a , F a )  | a in S }]]$. It is easy to verify that $[[R]]$ satisfy
the the constraint required in the precondition of \rref{I-Pi}.

@@

Unfortunately, we cannot define the function case of our logical
relation directly using \rref{I-PiAlt} since it not only violates the
syntactic strict positivity constraint required in proof assistants,
but is genuinely non-monotone when written as an endofunction over the
domain of relations.
Intuitively, the failure of monotonicity stems from the fact
that the witness picked in the precondition is not necessarily the
same witness being referred to in the post condition. While it might
be possible to restrict the domain with additional constraints such as
functionality and inversion properties, we opt for our current
formulation of \rref{I-Pi} so we immediately obtain a
well-defined inductive relation and a usable induction principle. The
slight disadvantage of \rref{I-Pi} is that we need to construct the
function $[[R]]$ each time we apply it, though this is mitigated by
the immediate admissibility of \rref{I-PiAlt}.
